General constraints on magic distillation protocols from statistical mechanics

### 2021-05-23 magic_paper.tex
98 percent happy with appendices. Couple of minor things to check/agree on.

We approach this problem using tools from statistical mechanics and recent work from the overlap of thermodynamics and quantum information theory~\cite{cit:horodecki2, Brandao_2015, Vinjanampathy_2016, Goold_2016, cit:lostaglio, cit:gour}\nick{Can't find Danzig's paper anywhere}. In particular, we shall view \emph{magic as a form of free energy}, relative to the set of stabilizer states. This perspective takes the convex set of stabilizer states as the set of ``equilibrium'' states, and magic corresponds to a form of non-equilibrium free energy that is strictly non-classical.

@@

\begin{figure}[t]
    \centering
        \includegraphics[scale=0.4]{figs/protocol_diagram.pdf}
    \caption{\textbf{Analysing magic protocols relative to a stabilizer reference state.} 
	A magic protocol aims to turn $n$ copies of some magic state $\rho$ to $m$ copies of some other state $\rho'$.
	Constraints on the efficiency of the protocol can be imposed by considering its action on a separate probe state $\tau$.
    }
    \label{fig:sketch}
\end{figure}

@@

The above bound depends on: (a) quantum computational measures $\epsilon, \epsilon'$, (b) thermodynamic quantities $F, F',$  and (c) intermediate terms $\phi, \phi'$. The intermediate terms depend on how the energy eigenbasis of the system $\{|E_k\>\}$ relates to the computational stabilizer basis $\{|k\>\}$.  In particular, the energetic term $\phi$ quantifies the degree to which the negativity in the Wigner representation can be ascribed a sharp energetic value for the given Hamiltonian of the system. Its form is similar to $F$, and so can informally be viewed as a `magic free energy' term, associated to a `magic partition function' $\zeta$. The coefficients $\alpha_k$ may be negative for some $k$, when the Hamiltonian eigen-decomposes into non-stabilizer states, but the quantity $\phi$ is always well-defined, since the function $\zeta$ is a positive multiple of the stabilizer Wigner component $W_\tau(0,0)$, as seen in the proof of the theorem. In the case that the system's Hamiltonian $H$ has an eigenbasis consisting of stabilizer states then we show in Appendix \ref{sharp-phi} that $\phi$ takes on a sharp, temperature-independent value $\phi \in \{E_0,E_1,E_2\}$.

@@

@@

\begingroup
\def\thetheorem{\ref{lem:lcmax}}
\begin{theorem}
	Given a magic state $\rho$, the maximum $L_\star$ of its Lorenz curve $\lc{\rho}{\sigma}(x)$ is independent of the $\sigma$--fragment and equal to $1+\sn{\rho}$. Moreover, the majorization constraint is stronger than mana in every fragment.
\end{theorem}
\addtocounter{theorem}{-1}
\endgroup
\begin{proof}
	We denote the Wigner distributions of the states single-component vectors $\w(\rho)=W_\rho(x)$ and $ \w(\sigma)=W_\sigma(x)$. Likewise, we write $\w(\rho|\sigma) = W_{\rho|\tau}(\x)$.
	We choose the component indexing so that $\w(\rho|\sigma)^\downarrow = \w(\rho|\sigma)$ and so the components are sorted in non-increasing order.

@@

Then, given any coordinates $(x, L)$ and $(x', L')$ of the input and output Lorenz curves respectively, where $0 < x \leq X_0$ and $0 < x' \leq X'_0$, the process is possible only if
\begin{equation}\label{eq:first_elb_bound1}
	\frac{L}{x} \geq \frac{L'}{x'}.
\end{equation}
\end{proposition}
\begin{proof}
Since both pairs of coordinates are located between $(0,0)$ and the first elbow of their respective curves, we can derive the bound via a simple interpolation on the line segment connecting the origin and the appropriate pair of coordinates.

@@

We now show that the relative majorization condition can be reduced to considering the elbows of the Lorenz curves.
\begin{proposition}\label{thm:elbows}
	Let $\rho, \rho'$ be two quantum states with Lorenz curves $\lc{\rho}{\sigma}(x), \lc{\rho'}{\sigma'}(x)$ in the $\sigma$--fragment.

	Let $t$ be the number of elbows of $\lc{\rho'}{\sigma'}(x)$ at locations $x_1, \dots, x_t$.

	Then, $\lc{\rho}{\sigma}(x) \geq \lc{\rho'}{\sigma'}(x)$ for all $x \in [0,1]$ iff $\lc{\rho}{\sigma}(x_{i}) \geq \lc{\rho'}{\sigma'}(x_{i})$ for all $i =1,\dots,t$.
\end{proposition}
\begin{proof}	
	$\lc{\rho}{\sigma}(x) \geq \lc{\rho'}{\sigma'}(x)$ for all $x \in [0,1]$ trivially implies $\lc{\rho}{\sigma}(x_{i}) \geq \lc{\rho'}{\sigma'}(x_{i})$ for all $i = 1,\dots,n'$.

@@

	Due to concavity of $\lc{\rho}{\sigma}$, it is clear that for all $x \in [x_{i}, x_{i+1}]$, we have $\lc{\rho}{\sigma}(x) \geq \ell_\rho(x) \geq \ell_\rho'(x) = \lc{\rho'}{\sigma}(x)$.
	This argument can be made in all intervals $[x_{i}, x_{i+1}]$ with $i=0,\dots,n'$, so the proof is complete.
\end{proof}
The above theorem can be used in reducing the necessary distillation constraints derived via majorization in $\sigma$--fragments.

Any free state $\sigma \in \F$ corresponds to a $d^2$--dimensional probability distribution $\W{\sigma}$ and any free operation $\E \in \O$ corresponds to a $d^2 \times d^2$ stochastic matrix (or conditional probability distribution) $\W{\E}$.
Note that these mappings are one-to-one due to the orthogonality of the phase-point operators as an operator basis.
Note further that free states $\F$ are mapped onto a \emph{strict subset} of the set of probability distributions.
As a counterexample, the sharp $d^2$--dimensional probability distribution $(1, 0, \dots, 0)$ does not correspond to any qudit Wigner distribution because of the boundedness condition in~\cref{thm:wstate}.
Similarly, not all stochastic matrices correspond to completely positive operations.
As an example, consider the permutation matrix
\begin{equation}
    \Pi_X = \begin{psmallmatrix}
        0 & 1 & 0 & 0 & 0 \\
        0 & 0 & 0 & 0 & 1 \\
        0 & 0 & 0 & 1 & 0 \\
        1 & 0 & 0 & 0 & 0 \\
        0 & 0 & 1 & 0 & 0
    \end{psmallmatrix} \otimes \begin{psmallmatrix}
        0 & 0 & 1 & 0 & 0 \\
        0 & 0 & 0 & 0 & 1 \\
        0 & 0 & 0 & 1 & 0 \\
        1 & 0 & 0 & 0 & 0 \\
        0 & 1 & 0 & 0 & 0    
    \end{psmallmatrix} \in {\rm{S}}_5({\W{\frac{1}{5}\id}}).
\end{equation}
It preserves the uniform distribution $\W{\frac{1}{5}\id}$, but it does not correspond to any positive (hence quantum) operation.

@@

\subsection{Binomial distributions and error bounds}\label{app:phi}
Consider an experiment consisting of $n$ trials of throwing a $p$--coin, that is a coin with probability $p$ of landing on one side and $1-p$ of landing on the other.
We denote the sums over an even number $m$ of successful trials by $\Phi_+(m; n, p)$ and an odd number $m$ of successful trials by $\Phi_-(m; n, p)$, and they are given by
\begin{align}	
	\Phi_+(m; n, p) &\coloneqq \sum\limits_{\ell=0}^{m/2} \binom{n}{2\ell} p^{2\ell} (1-p)^{n-2\ell}, \nonumber\\ 
	&\text{for even integers } m\in[0,n], \label{eq:fp_app} \\
	\Phi_-(m; n, p) &\coloneqq \sum\limits_{\ell=1}^{(m-1)/2} \binom{n}{2\ell+1} p^{2\ell+1} (1-p)^{n-(2\ell+1)}, \nonumber\\ 
	&\text{for odd integers }m\in[0,n]. \label{eq:fn_app}
\end{align}
Note that index $m$ only takes even (odd) values when labelling $\Phi_+$ ($\Phi_-$).
In~\cref{app:lcsu_coord}, we will use $\Phi_+$ and $\Phi_-$ to express the elbow coordinates of Lorenz curves in the unital fragment.

We also define the classical entropy of a $p$--coin and the classical relative entropy between a $p$--coin and a $q$--coin,
\begin{align}
	S(p) &\coloneqq -p\log{p} -(1-p)\log{(1-p)}, \label{eq:ent}\\
	\ent{p}{q} &\coloneqq p \log{\frac{p}{q}} + (1-p) \log{\frac{1-p}{1-q}}. \label{eq:ent_rel}
\end{align}
They are symmetric about $p=1/2$, namely $S(p) = S(1-p)$ and $\ent{p}{q} = \ent{1-p}{1-q}$.

A useful result is the entropic bound on a combination~\cite{cit:ash}.
\begin{lemma}\label{lem:comb_bounds}
	For all $\ell\in [1,n-1]$,
	\begin{align}
		&\left[ 8\ell\left(1-\frac{\ell}{n}\right) \right]^{-\frac{1}{2}} 2^{n S\left(\frac{\ell}{n}\right)} \leq \binom{n}{\ell} \leq \\
		&\left[ 2\pi \ell\left(1-\frac{\ell}{n}\right) \right]^{-\frac{1}{2}} 2^{n S\left(\frac{\ell}{n}\right)}.
	\end{align}
\end{lemma}
The proof provided in~\cite{cit:ash} proceeds with direct calculation for the edge cases $\ell = 1,2, n-1, n-2$ and use Stirling's approximation for the remaining cases.

We can use~\cref{lem:comb_bounds} to provide strict upper and lower bounds on the functions $\Phi_+, \Phi_-$.

\subsection{Theory on bounding the core functions}
Here we present a more manageable method of bounding the core functions $\Phi_{\pm}$, which however results in looser bounds. 
We can rewrite the functions as
\begin{equation}
	\Phi_{\pm}(m; n, a) = \frac{1}{2}(\Phi(m; n, a) \pm (1+a)^{-n} S(m; n, a)),
\end{equation}
where we have substituted $a = p/(1-p)$. 
$\Phi$ is the standard cumulative function
\begin{equation}
	\Phi(m; n, a) = (1+a)^{-n} \sum_{k=0}^m \binom{n}{k} a^k,
\end{equation}
and the remainder term is
\begin{equation}
	S(m; n, a) \coloneqq \sum_{k=0}^m \binom{n}{k} (-a)^k.
\end{equation}
We have the following asymptotic bounds on the behaviour of $\Phi$~\cite{cit:ash},
\begin{lemma}\label{lem:phi_bounds}
	Given fixed $n>0$ and $p$, $\Phi$ satisfies the following bounds:
	\begin{align*}
		\begin{split}
		&\text{1. } \Phi(m; n, p) \geq \left[ 8m\left(1-\frac{m}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ent{\frac{m}{n}}{p}}, \\
		&\hspace{14pt} m\in [1,n-1] \\
		&\text{2. } \Phi(m; n, p) \geq 1 - 2^{-n\ent{\frac{m+1}{n}}{p}},\ m\in [np+1,n-2] \\
		&\text{3. } \Phi(m; n, p) \leq 1 - \left[ 8(m+1)\left(1-\frac{m+1}{n}\right) \right]^{-\frac{1}{2}}\times \\
		&\hspace{14pt} 2^{-n\ent{\frac{m+1}{n}}{p}},\ m\in [0,n-2]
		\end{split}
		\\
		&\text{4. } \Phi(m; n, p) \leq 2^{-n\ent{\frac{m}{n}}{p}},\ m\in [0,np]
	\end{align*}
\end{lemma}

We would like some theory that estimates the value of $S(m; n, a)$ for different parameter regimes. 
We can consider the function $f(x) = (1+x)^n$ and note that $S(m; n, a)$ is the $m$'th partial sum of this expansion at the point $x=-a$.

The truncated Maclaurin series of a general function $f(x)$ is
\begin{equation}
	f(x) = f(0) + x f'(0) + \dots \frac{x^m}{m!}f^{(m)}(0) + R_m(x)
\end{equation}
with a remainder term
\begin{align}
	R_m (x)&= \int_{0}^x dt f^{(m+1)}(t) \frac{(x-t)^m}{m!} \\
	&= \frac{x^{m+1}}{(m+1)!} f^{(m+1)}(x_*),
\end{align}
where in the second expression, $x_*$ is an implicit point that lies between $0$ and $x$ that comes from the Mean Value Theorem.

Applying this to the function $f(x) = (1+x)^n$ gives
\begin{equation}
	(1+x)^n = \sum_{k=0}^m \binom{n}{k} x^k + R_m.
\end{equation}
Evaluating at $x=-a$ gives
\begin{equation}
	S(m; n, a) = (1-a)^n - R_m(-a),
\end{equation}
where the key remainder term is given by
\begin{align}
	R_m(-a) &= \int_0^{-a} dt f^{(m+1)}(t) \frac{(-a-t)^m}{m!} \\
&= \frac{(-a)^{m+1}}{(m+1)!} f^{(m+1)}(x_*).
\end{align}
We can also compute the derivative $f^{(m+1)}(x)$ explicitly,
\begin{equation}
	f^{(m+1)}(x) = (m+1)!\binom{n}{m+1}(1+x)^{n-m-1}.
\end{equation}
Therefore, we have that
\begin{align}
	R_m(-a) &= (-1)^{m+1}(m+1)\binom{n}{m+1}\times \nonumber\\
	&\hspace{12pt} \int_{-a}^0 dt (1+t)^{n-m-1}(a+t)^m \\
&= \binom{n}{m+1}(-a)^{m+1}(1+x_*)^{n-m-1},
\end{align}
where in the latter expression $x_* \in [-a,0]$. 
Note that the first integral expression can be estimated via the Cauchy-Schwarz or the H{\"o}lder inequality. 
Therefore, we can either work with an explicit form with an unknown (but bounded) parameter $x_*$, or we can use the integral form and provide concrete estimates on it value.

A very simple estimate, based on $x_*$ lying in the interval $[-a,0]$ gives
\begin{equation}
(1-a)^{n-m-1} \leq \frac{R_m(-a)}{\binom{n}{m+1}(-a)^{m+1}} \leq 1,
\end{equation}
which in turn leads to the following bounds on $\Phi_+(m; n, a)$:
\begin{align}
	\hspace{-2cm}2\Phi_+(m; n, a) \leq\ &\Phi(m; n, a) + (1-a)^n - \frac{(-a)^{m+1}}{m!} \times \nonumber\\
	 &\binom{n}{m+1} (1-a)^{n-m-1} \text{ and} \\
	2\Phi_+(m; n, a) \geq\ &\Phi(m; n, a) + (1-a)^n - \frac{(-a)^{m+1}}{m!}  \times \nonumber\\
	 &\binom{n}{m+1} .
\end{align}

The Wigner distribution of the $n$--copy $\epsilon$--noisy Strange state $\rho_{\rm{S}}(\epsilon)^{\otimes n}$ is given by $\W{\rho_{\rm{S}}(\epsilon)^{\otimes n}} = W_{\rho_{\rm{S}}(\epsilon)}^{\otimes n}$ and, using~\cref{lem:ncopycomponents} in the unital fragment, we can associate it with the set $\{(w_i, m_i)\}_{i=0,\dots,n}$ of component-multiplicity pairs, as presented in~\cref{tab:lcsu} for all different combinations of parameters $n,\epsilon$.
We split the values of index $i$ into two intervals, defined precisely as
\begin{align}
&\text{LHS: } 0 \leq i \leq \left\lfloor \frac{n}{2} \right\rfloor \text{ and} \\
&\text{RHS: } \left\lfloor \frac{n}{2} \right\rfloor +1 \leq i \leq n.
\end{align}
The labels $\text{LHS}, \text{RHS}$ imply that the components contribute to the left hand side or the right hand side of the maximum (including the maximum) of the Lorenz curve.
To compactify notation, we also use multiplication modulo $(n+1)$, specifically, for any positive integer $m \leq 2n$:
\begin{equation}
	m \hspace{-6pt}\mod{\hspace{-2pt}(n+1)} = 
	\begin{cases}
		m,\ &\text{if } 0 \leq m \leq n, \\
		m - n - 1,\ &\text{if } n < m \leq 2n.
	\end{cases}
\end{equation}
\begin{table}[h]
  \def\arraystretch{1.5}
  \centering
  \begin{tabular}{c|c|c|r|r}
    \multicolumn{3}{c|}{Case} & \multicolumn{1}{c}{$m_{i}$} & \multicolumn{1}{|c}{$w_{i}$} \\[0.5ex]\hline
    \multirow{4}{*}{\raisebox{-4ex}{\rotatebox[origin=c]{90}{$0\leq \epsilon < \frac{3}{7}$}}} & \hspace{0.8ex}\multirow{2}{*}{\raisebox{-1ex}{\rotatebox[origin=c]{90}{$n$ even}}}\hspace{0.8ex} & LHS & $8^{2i}\binom{n}{2i}$ & $u^{2i}v^{n-2i}$ \\
    & & RHS & $8^{n-2i}\binom{n}{2i}$ & $-u^{n-2i}v^{2i}$ \\ \cline{2-5}
    & \multirow{2}{*}{\raisebox{-2ex}{\rotatebox[origin=c]{90}{$n$ odd}}} & LHS & $8^{2i+1}\binom{n}{2i+1}$ & $u^{2i+1}v^{n-2i-1}$ \\
    & & RHS & $8^{n-2i-1}\binom{n}{2i+1}$ & $-u^{n-2i-1}v^{2i+1}$ \\ \hline
    \multirow{4}{*}{\raisebox{-4ex}{\rotatebox[origin=c]{90}{$\frac{3}{7}\leq \epsilon < \frac{3}{4}$}}} & \multirow{2}{*}{\raisebox{-1ex}{\rotatebox[origin=c]{90}{$n$ even}}} & LHS & $8^{n-2i}\binom{n}{2i}$ & $u^{n-2i}v^{2i}$ \\
    & & RHS & $8^{2i}\binom{n}{2i}$ & $-u^{2i}v^{n-2i}$ \\ \cline{2-5}
    & \multirow{2}{*}{\raisebox{-2ex}{\rotatebox[origin=c]{90}{$n$ odd}}} & LHS & $8^{n-2i}\binom{n}{2i}$ & $u^{n-2i}v^{2i}$ \\
    & & RHS & $8^{2i}\binom{n}{2i}$ & $-u^{2i}v^{n-2i}$ \\ \hline
  \end{tabular}
  \caption{Wigner components $w_{i}(n, \epsilon)$ of $\rho_{\rm{S}}(\epsilon)^{\otimes n}$ along with their multiplicities $m_{i}(n, \epsilon)$, with $0 \leq i \leq n$.
  The expressions change depending on the error $\epsilon$, the parity of the number of copies $n$ and whether the index $i$ is lower or higher than the index of the Lorenz curve maximum (LHS vs RHS).
  Multiplication $2i$ is considered $\hspace{-6pt}\mod{\hspace{-2pt}(n+1)}$.}
  \label{tab:lcsu}
\end{table}

>>>>

We approach this problem using tools from statistical mechanics and recent work from the overlap of thermodynamics and quantum information theory~\cite{cit:horodecki2, Brandao_2015, Vinjanampathy_2016, Goold_2016, cit:lostaglio, cit:gour}\nick{Can't find Danzig's paper anywhere}\ddd{[Remembered name incorrectly, is this one: Janzing, Dominik, et al. "Thermodynamic cost of reliability and low temperatures: tightening Landauer's principle and the second law." Put this one first in list -- people always miss it, but it is the first.] }. In particular, we shall view \emph{magic as a form of free energy}, relative to the set of stabilizer states. This perspective takes the convex set of stabilizer states as the set of ``equilibrium'' states, and magic corresponds to a form of non-equilibrium free energy that is strictly non-classical.

@@

\begin{figure}[t]
    \centering
        \includegraphics[scale=0.4]{figs/protocol_diagram.pdf}
    \caption{\textbf{Analysing magic protocols relative to a stabilizer reference state.} 
	A magic protocol aims to turn $n$ copies of some magic state $\rho$ to $m$ copies of some other state $\rho'$.
	Constraints on the efficiency of the protocol can be imposed by considering its action on a separate probe state $\tau$.
    } \ddd{[Not the clearest - we can zoom-chat on this.]}
    \label{fig:sketch}
\end{figure}

@@

The above bound depends on: (a) quantum computational measures $\epsilon, \epsilon'$, (b) thermodynamic quantities $F, F',$  and (c) intermediate terms $\phi, \phi'$. The intermediate terms depend on how the energy eigenbasis of the system $\{|E_k\>\}$ relates to the computational stabilizer basis $\{|k\>\}$.  In particular, the energetic term $\phi$ quantifies the degree to which the negativity in the Wigner representation can be ascribed a sharp energetic value for the given Hamiltonian of the system. Its form is similar to $F$, and so can informally be viewed as a `magic free energy' term, associated to what can informally be viewed as a `partition function' $\zeta$. The coefficients $\alpha_k$ may be negative for some $k$, when the Hamiltonian eigen-decomposes into non-stabilizer states, but the quantity $\phi$ is always well-defined, since the function $\zeta$ is a positive multiple of the stabilizer Wigner component $W_\tau(0,0)$, as seen in the proof of the theorem. In the case that the system's Hamiltonian $H$ has an eigenbasis consisting of stabilizer states then we show in Appendix \ref{sharp-phi} that $\phi$ takes on a sharp, temperature-independent energy value $\phi \in \{E_0,E_1,E_2\}$.

@@

\begin{lemma}\label{L-concave} Let $\p$ be a quasi-probability distribution and let $\r$ be a probability distribution with strictly non-zero components. Then $L_{\p|\r}(x)$ is a concave function on $[0,1]$.
\begin{proof}Let $L_{\p|\r}(x)$ attain its maximum at $x=x_\star$. Then from the definition, the function is continuous, monotone increasing on $[0,x_\star]$, and monotone decreasing on $[x_\star, 1]$. Consider $a,b \in [0,1]$ with $a <b$ and some probability $q \in [0,1]$. If $a$ and $b$ both lie in either the interval $[0,x_\star]$ or $[x_\star, 1]$ then concavity follows immediately. Otherwise $a$ is in the first interval and $b$ in the second interval. If $w=qa + (1-q)b$ lies in the first interval then
\begin{align}
qL_{\p|\r}(a) + (1-q)L_{\p|\r}(b) &\le qL_{\p|\r}(a) + (1-q)L_{\p|\r}(x_\star)\nonumber\\
& \le L_{\p|\r}(w)
\end{align}
from monotonicity on the first interval. Otherwise $w$ lies in the second interval and
\begin{align}
qL_{\p|\r}(a) + (1-q)L_{\p|\r}(b) &\le qL_{\p|\r}(x_\star) + (1-q)L_{\p|\r}(b)\nonumber\\
& \le L_{\p|\r}(w)
\end{align}
which follows from monotonicity on the second interval. Therefore the Lorenz curve is concave on $[0,1]$.
\end{proof}
\end{lemma}

@@

\begingroup
\def\thetheorem{\ref{lem:lcmax}}
\begin{theorem}
	Given a magic state $\rho$, the maximum value $L_\star$ attained by the Lorenz curve $\lc{\rho}{\sigma}(x)$ is independent of the $\sigma$--fragment and equal to $1+\sn{\rho}$. Moreover, the relative majorization constraint is stronger than mana in every fragment.
\end{theorem}
\addtocounter{theorem}{-1}
\endgroup
\begin{proof}
	We denote the Wigner distributions of the states single-component vectors $\w(\rho)=W_\rho(x)$ and $ \w(\sigma)=W_\sigma(x)$. Likewise, we write $\w(\rho|\sigma) = W_{\rho|\tau}(\x)$.
	We choose the component indexing so that $\w(\rho|\sigma)^\downarrow = \w(\rho|\sigma)$ and so the components are sorted in non-increasing order.

@@

Then, given any coordinates $(x, L)$ and $(x', L')$ on the input and output Lorenz curves respectively, where $0 < x \leq X_0$ and $0 < x' \leq X'_0$, the process is possible only if
\begin{equation}\label{eq:first_elb_bound1}
	\frac{L}{x} \geq \frac{L'}{x'}.
\end{equation}
\end{proposition}
\ddd{[I think the following is a more direct/informative proof]}
\ddd{\begin{proof} If the transformation is possible then $L_{\rho|\sigma}(x) \ge L_{\rho'|\sigma'} (x)$ for all $x$ in $[0,1]$. If we restrict to the initial linear region segment of $L_{\rho|\sigma}(x)$ joining $(0,0)$ to $(X_0,L_0)$ then we require this to be above the initial linear segment of $L_{\rho'|\sigma'}(x)$ joining $(0,0)$ to $(X_0', L_0')$. Since both these linear segments go through the origin this condition implies that the slope of the initial segment of the input Lorenz curve must be greater than or equal to the slope of the initial segment of the output Lorenz curve. This slope constraint is then equivalent to Equation \ref{eq:first_elb_bound1}. \end{proof}}
\begin{proof}
Since both pairs of coordinates are located between $(0,0)$ and the first elbow of their respective curves, we can derive the bound via a simple interpolation on the line segment connecting the origin and the appropriate pair of coordinates.

@@

We now show that the relative majorization condition can be reduced to considering the elbows of the output Lorenz curve.
\begin{proposition}\label{thm:elbows}
	Let $\rho, \rho'$ be any two quantum states with Lorenz curves $\lc{\rho}{\sigma}(x), \lc{\rho'}{\sigma'}(x)$, where $\sigma, \sigma'$ are states with positive Wigner distributions. Assume that $\lc{\rho'}{\sigma'}(x)$ has $t$ elbows at locations $x_1, \dots, x_t$. Then, $\lc{\rho}{\sigma}(x) \geq \lc{\rho'}{\sigma'}(x)$ for all $x \in [0,1]$ iff $\lc{\rho}{\sigma}(x_{i}) \geq \lc{\rho'}{\sigma'}(x_{i})$ for all $i =1,\dots,t$.
\end{proposition}
\begin{proof}	
If $\lc{\rho}{\sigma}(x) \geq \lc{\rho'}{\sigma'}(x)$ for all $x \in [0,1]$ then the condition on the elbows trivially follows.

\ddd{Conversely, assume $L_{\rho|\sigma}(x_i) \ge L_{\rho'|\sigma'}(x_i)$ for all $i=1,\dots t$. Suppose on the contrary that $L_{\rho|\sigma}(x)$ dips below $L_{\rho'|\sigma'}(x)$ at some point $y$ where $x_i < y < x_{i+1}$ for some $i$. Then this implies
\begin{align}
L_{\rho|\sigma}(x_i) &\ge L_{\rho'|\sigma'}(x_i) \nonumber \\
L_{\rho|\sigma}(y) & <L_{\rho'|\sigma'}(y) \nonumber \\
L_{\rho|\sigma}(x_{i+1}) &\ge L_{\rho'|\sigma'}(x_{i+1}).
\end{align}
However since $L_{\rho'|\sigma'}(x)$ is linear between $x_i$ and $x_{i+1}$, the above conditions imply that $L_{\rho|\sigma}(x)$ is not a concave function, which contradicts Lemma \ref{L-concave}. Therefore $L_{\rho|\sigma}(x) \ge L_{\rho'|\sigma'}(x)$ for all $x$ in $[0,1]$.
}

@@

	Due to concavity of $\lc{\rho}{\sigma}$, it is clear that for all $x \in [x_{i}, x_{i+1}]$, we have $\lc{\rho}{\sigma}(x) \geq \ell_\rho(x) \geq \ell_\rho'(x) = \lc{\rho'}{\sigma}(x)$.
	This argument can be made in all intervals $[x_{i}, x_{i+1}]$ with $i=0,\dots,n'$, so the proof is complete.
\end{proof}
The above result simplifies the computation of relative majorization constraints to looking at the elbows of the output Lorenz curve.

@@

\subsection{Binomial distributions and error bounds}\label{app:phi}
Consider an experiment consisting of $n$ trials of throwing a $p$--coin, that is a coin with probability $p$ of landing on one side and $1-p$ of landing on the other.
We denote the sums over an even number $m$ of successful trials by $\Phi_+(m; n, p)$ and an odd number $m$ of successful trials by $\Phi_-(m; n, p)$, and they are given by
\begin{align}	
	\Phi_+(m; n, p) &\coloneqq \sum\limits_{\ell=0}^{m/2} \binom{n}{2\ell} p^{2\ell} (1-p)^{n-2\ell}, \nonumber\\ 
	&\text{for even integers } m\in[0,n], \label{eq:fp_app} \\
	\Phi_-(m; n, p) &\coloneqq \sum\limits_{\ell=1}^{(m-1)/2} \binom{n}{2\ell+1} p^{2\ell+1} (1-p)^{n-(2\ell+1)}, \nonumber\\ 
	&\text{for odd integers }m\in[0,n]. \label{eq:fn_app}
\end{align}
In~\cref{app:lcsu_coord}, we will use $\Phi_+$ and $\Phi_-$ to express the elbow coordinates of Lorenz curves in the unital fragment.

We have the Shannon entropy of a $p$--coin and the Shannon relative entropy for a $p$--coin and a $q$--coin given by the expressions
\begin{align}
	S(p) &\coloneqq -p\log{p} -(1-p)\log{(1-p)}, \label{eq:ent}\\
	\ent{p}{q} &\coloneqq p \log{\frac{p}{q}} + (1-p) \log{\frac{1-p}{1-q}}. \label{eq:ent_rel}
\end{align}
They are symmetric about $p=1/2$, namely $S(p) = S(1-p)$ and $\ent{p}{q} = \ent{1-p}{1-q}$. We also have the entropic bounds
\begin{lemma}\label{lem:comb_bounds}
	For all $\ell\in [1,n-1]$,
	\begin{align}
		&\left[ 8\ell\left(1-\frac{\ell}{n}\right) \right]^{-\frac{1}{2}} 2^{n S\left(\frac{\ell}{n}\right)} \leq \binom{n}{\ell} \leq \\
		&\left[ 2\pi \ell\left(1-\frac{\ell}{n}\right) \right]^{-\frac{1}{2}} 2^{n S\left(\frac{\ell}{n}\right)}.
	\end{align}
\end{lemma}
The proof provided in~\cite{cit:ash} proceeds with direct calculation for the edge cases $\ell = 1,2, n-1, n-2$ and use Stirling's approximation for the remaining cases. The bounds~\cref{lem:comb_bounds} can be used to derive the following upper and lower bounds on the functions $\Phi_+, \Phi_-$.

We rewrite the functions as
\begin{equation}
	\Phi_{\pm}(m; n, a) = \frac{1}{2}(\Phi(m; n, a) \pm (1+a)^{-n} S(m; n, a)),
\end{equation}
where $a = p/(1-p)$, and $\Phi$ is the standard cumulative function for the binomial distribution
\begin{equation}
	\Phi(m; n, a) = (1+a)^{-n} \sum_{k=0}^m \binom{n}{k} a^k,
\end{equation}
and we define the remainder term
\begin{equation}
	S(m; n, a) \coloneqq \sum_{k=0}^m \binom{n}{k} (-a)^k.
\end{equation}
The cumulative function $\Phi$ obeys the following bounds, which are proven in~\cite{cit:ash}.
\begin{lemma}\label{lem:phi_bounds}
	Given fixed $n>0$ and $p$, $\Phi$ satisfies the following bounds:
	\begin{align*}
		\begin{split}
		&\text{1. } \Phi(m; n, p) \geq \left[ 8m\left(1-\frac{m}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ent{\frac{m}{n}}{p}}, \\
		&\hspace{14pt} m\in [1,n-1] \\
		&\text{2. } \Phi(m; n, p) \geq 1 - 2^{-n\ent{\frac{m+1}{n}}{p}},\ m\in [np+1,n-2] \\
		&\text{3. } \Phi(m; n, p) \leq 1 - \left[ 8(m+1)\left(1-\frac{m+1}{n}\right) \right]^{-\frac{1}{2}}\times \\
		&\hspace{14pt} 2^{-n\ent{\frac{m+1}{n}}{p}},\ m\in [0,n-2]
		\end{split}
		\\
		&\text{4. } \Phi(m; n, p) \leq 2^{-n\ent{\frac{m}{n}}{p}},\ m\in [0,np]
	\end{align*}
\end{lemma}

We can now put estimates on $S(m; n, a)$, and hence the functions $\Phi_{\pm}$, for different parameter regimes. 
We can consider the function $f(x) = (1+x)^n$ and note that $S(m; n, a)$ is the $m$'th partial sum of this expansion at the point $x=-a$. The truncated Maclaurin series is given by
\begin{equation}
	f(x) = f(0) + x f'(0) + \dots \frac{x^m}{m!}f^{(m)}(0) + R_m(x)
\end{equation}
with a remainder term
\begin{align}
	R_m (x)&= \int_{0}^x dt f^{(m+1)}(t) \frac{(x-t)^m}{m!} \\
	&= \frac{x^{m+1}}{(m+1)!} f^{(m+1)}(x_*),
\end{align}
where in the second expression, $x_*$ is a point that lies between $0$ and $x$ that comes from the Mean Value Theorem.

Applying this to the function $f(x) = (1+x)^n$ gives
\begin{equation}
	(1+x)^n = \sum_{k=0}^m \binom{n}{k} x^k + R_m.
\end{equation}
Therefore if we evaluate at $x=-a$ we obtain
\begin{equation}
	S(m; n, a) = (1-a)^n - R_m(-a),
\end{equation}
where the remainder term is given by
\begin{align}
	R_m(-a) &= \int_0^{-a} dt f^{(m+1)}(t) \frac{(-a-t)^m}{m!} \\
&= \frac{(-a)^{m+1}}{(m+1)!} f^{(m+1)}(x_*).
\end{align}
We can compute the derivative $f^{(m+1)}(x)$ explicitly,
\begin{equation}
	f^{(m+1)}(x) = (m+1)!\binom{n}{m+1}(1+x)^{n-m-1}.
\end{equation}
Therefore, we have that
\begin{align}
	R_m(-a) &= (-1)^{m+1}(m+1)\binom{n}{m+1}\times \nonumber\\
	&\hspace{12pt} \int_{-a}^0 dt (1+t)^{n-m-1}(a+t)^m \\
&= \binom{n}{m+1}(-a)^{m+1}(1+x_*)^{n-m-1},
\end{align}
where in the latter expression $x_* \in [-a,0]$. 
The first integral expression can be estimated via the Cauchy-Schwarz or the H{\"o}lder inequality. 
Therefore, one has an explicit form with an unknown (but bounded) parameter $x_*$, or the integral form.

A very simple estimate, based on $x_*$ lying in the interval $[-a,0]$ gives
\begin{equation}
(1-a)^{n-m-1} \leq \frac{R_m(-a)}{\binom{n}{m+1}(-a)^{m+1}} \leq 1,
\end{equation}
which in turn leads to the following bounds on $\Phi_+(m; n, a)$:
\begin{align}
	\hspace{-2cm}2\Phi_+(m; n, a) \leq\ &\Phi(m; n, a) + (1-a)^n - \frac{(-a)^{m+1}}{m!} \times \nonumber\\
	 &\binom{n}{m+1} (1-a)^{n-m-1} \text{ and} \\
	2\Phi_+(m; n, a) \geq\ &\Phi(m; n, a) + (1-a)^n - \frac{(-a)^{m+1}}{m!}  \times \nonumber\\
	 &\binom{n}{m+1} .
\end{align}
Better bounds can be obtained with a finer analysis, and it would be of interest to obtain tighter asymptotic behaviour on $\Phi_{\pm}$ that could in turn be used to improve magic distillation bounds.
